\textsf{Are large language models superhuman chemists?}

### 2024-10-18 src/tex/ms.tex
add more details

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that all humans answered.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions). 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
We then computed the percentage of questions the models or humans answered correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

\paragraph{Participants}
% Users were open to reporting about their experience in chemistry. 
% Overall, \variable{output/num_users_with_education_info.txt} did so. 
% Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
% \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.

>>>>

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that all humans answered.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions). 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
We then computed the percentage of questions the models or humans answered correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}

@@

\paragraph{Participants}
Users were open to reporting about their experience in chemistry. 
Overall, \variable{output/num_users_with_education_info.txt} did so. 
Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
